KIN: A method to infer relatedness from low-coverage ancient DNA

### 2022-05-13 paper_overleaf/main.tex
Updates from Overleaf

\paragraph{IBD prediction}
We investigated the accuracy of IBD state prediction along the genome by counting the number of genomic windows where we correctly predict IBD state for different relatdness and coverages (Fig. \ref{fig4:IBDstate_accuracy}). The accuracy for coverages of 0.1x or higher is consistent and varies between 1.0 and 0.836, depending on relatedness. However, the accuracy decreases at lower coverages for most relatedness and ranges from 1.0 to 0.563 at 0.03x. On the other hand, the accuracy for identical individuals and parent-child pairs is nearly perfect for all investigated coverages, even at 0.03x coverage.
[add some comments about contamination and ascertainment bias, with reference to supplementary figures]

\begin{figure}[h!]
    \includegraphics[width=16cm]{plots/plotimg/plot_IBDaccuracy.png}
    \centering
    \caption{Evaluation of IBD estimation at different coverages. y-axis shows accuracy calculated over 60 simulations with each relatedness case and six different coverages shown on x-Here,Here,axis. The error bars are drawn at 1 standard deviation from the mean. Here, accuracy corresponding to relatedness cases for Parent-Child and Identical individuals is always 1, and overlaps with each other.}
    \label{fig4:IBDstate_accuracy}
\end{figure}

@@

\begin{table}
\caption{\label{tab:Table 2}Model performance for ROH prediction. Here we test ROH-HMM in four different cases of simulations: Control (without ROH, ascertainment bias or contamination, R (with ROH), RA (with ROH and ascertainment bias), RC (with ROH and contamination)}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
    Coverage & Control(bias) & Control(var) & R(bias) & R(var) & RA(bias) & RA(var) & RC(bias) & RC(var)\\
    \hline
    4x & -0.007 & 0.006 & -0.081 & 0.036 & -0.083 & 0.036 & -0.091 & 0.039\\
    \hline
    0.2x & -0.003 & 0.003 & -0.068 & 0.03 & -0.068 & 0.026 & -0.076 & 0.029\\
    \hline
    0.1x & -0.003 & 0.002 & -0.049 & 0.021 & -0.05 & 0.021 & -0.057 & 0.022\\
    \hline
\end{tabular}
\label{table1}
\end{table}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=18cm]{supplementary_info/plots/lcPlot.png}
    \caption{Comparison of IBD states estimated for Chagyrskaya specimens using (A) lcMLkin and (B) KIN. Each plot shows relatedness coefficient (calculated as $k_2+k_1/2$) on y-axis plotted against $k_0$ (proportion of genome with no chromosomes shared).
    The relatedness shown with different colors is estimated with both READ and KIN.}
    \label{figS6:Chagyrskaya_ibd}
\end{figure}

>>>>

\begin{table}[h!]{ l p{5cm}}
\caption{\label{tab:Table 2}Model performance for ROH prediction. Here we test ROH-HMM in four different cases of simulations: Control (without ROH, ascertainment bias or contamination, R (with ROH), RA (with ROH and ascertainment bias), RC (with ROH and contamination)}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
    Coverage & Control(bias) & Control(var) & R(bias) & R(var) & RA(bias) & RA(var) & RC(bias) & RC(var)\\
    \hline
    4x & -0.007 & 0.006 & -0.081 & 0.036 & -0.083 & 0.036 & -0.091 & 0.039\\
    \hline
    0.2x & -0.003 & 0.003 & -0.068 & 0.03 & -0.068 & 0.026 & -0.076 & 0.029\\
    \hline
    0.1x & -0.003 & 0.002 & -0.049 & 0.021 & -0.05 & 0.021 & -0.057 & 0.022\\
    \hline
\end{tabular}
\label{table1}
\end{table}

\paragraph{IBD prediction}
We investigated the accuracy of IBD state prediction along the genome by counting the number of genomic windows where we correctly predict IBD state for different relatdness and coverages (Fig. \ref{fig4:IBDstate_accuracy}). The accuracy for coverages of 0.1x or higher is consistent and varies between 1.0 and 0.836, depending on relatedness. However, the accuracy decreases at lower coverages for most relatedness and ranges from 1.0 to 0.563 at 0.03x. On the other hand, the accuracy for identical individuals and parent-child pairs is nearly perfect for all investigated coverages, even at 0.03x coverage. We find that adding contamination, ascertainment bias and ROH in our simulations does not affect IBD prediction for identical and parent-child pairs, but reduces accuracy for unrelated individuals by xx, yy and zz respectively, for second degree by xx,yy,zz and for siblings by xx,yy,zz respectively.

\begin{figure}[h!]
    \includegraphics[width=16cm]{plots/plotimg/plot_IBDaccuracy.png}
    \centering
    \caption{Evaluation of IBD estimation at different coverages. y-axis shows accuracy calculated over 60 simulations with each relatedness case and six different coverages shown on x-axis. The error bars are drawn at 1 standard deviation from the mean. Here, accuracy corresponding to relatedness cases for Parent-Child and Identical individuals is always 1, and overlaps with each other.}
    \label{fig4:IBDstate_accuracy}
\end{figure}

@@

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=18cm]{plots/plotimg/plot_IBDaccuracy_asc.png}
    \caption{Evaluation of IBD estimation in presence of ascertainment at different coverages. y-axis shows accuracy calculated over 60 simulations with each relatedness case and six different coverages shown on x-axis. The error bars are drawn at 1 standard deviation from the mean. Here, accuracy corresponding to relatedness cases for Parent-Child and Identical individuals is always 1, and overlaps with each other.}
    \label{figS11:ibd_asc}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=18cm]{plots/plotimg/plot_IBDaccuracy_roh.png}
    \caption{Evaluation of IBD estimation in presence of ROH at different coverages. y-axis shows accuracy calculated over 60 simulations with each relatedness case and six different coverages shown on x-axis. The error bars are drawn at 1 standard deviation from the mean. Here, accuracy corresponding to relatedness cases for Parent-Child and Identical individuals is always 1, and overlaps with each other.}
    \label{figS11:ibd_asc}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=18cm]{plots/plotimg/plot_IBDaccuracy_cont.png}
    \caption{Evaluation of IBD estimation in presence of contamination at different coverages. y-axis shows accuracy calculated over 60 simulations with each relatedness case and six different coverages shown on x-axis. The error bars are drawn at 1 standard deviation from the mean. Here, accuracy corresponding to relatedness cases for Parent-Child and Identical individuals is always 1, and overlaps with each other.}
    \label{figS11:ibd_asc}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=18cm]{supplementary_info/plots/lcPlot.png}
    \caption{Comparison of IBD states estimated for Chagyrskaya specimens using (A) lcMLkin and (B) KIN. Each plot shows relatedness coefficient (calculated as $k_2+k_1/2$) on y-axis plotted against $k_0$ (proportion of genome with no chromosomes shared).
    The relatedness shown with different colors is estimated with both READ and KIN.}
    \label{figS6:Chagyrskaya_ibd}
\end{figure}
